On exponential sums arising from the classical doubling method

### 2025-03-24 paper.tex
Jacobi sums modification

\newcommand{\GaussSumScalar}[2]{\mathrm{G}\left(#1, #2\right)}
\newcommand{\dblJacobiSumScalar}[2]{\mathrm{J}_{\pm}^{\mathrm{dbl}}\left(#1, #2\right)}
\newcommand{\posDblJacobiSumScalar}[2]{\mathrm{J}_{+}^{\mathrm{dbl}}\left(#1, #2\right)}
\newcommand{\negDblJacobiSumScalar}[2]{\mathrm{J}_{-}^{\mathrm{dbl}}\left(#1, #2\right)}
\newcommand{\dblVirtualJacobiSumScalar}[2]{\mathrm{j}_{\pm}^{\mathrm{dbl}}\left(#1, #2\right)}
\newcommand{\posDblVirtualJacobiSumScalar}[2]{\mathrm{j}_{+}^{\mathrm{dbl}}\left(#1, #2\right)}
\newcommand{\dblGammaFactor}[3]{\Gamma^{\mathrm{dbl}}\left(#1 \times #2, #3\right)}
\newcommand{\dblGammaFactorSpace}[4]{\Gamma^{\mathrm{dbl}}_{#1}\left(#2 \times #3, #4\right)}
\newcommand{\dblLanglandsGammaFactorSpace}[4]{\gamma^{\mathrm{dbl}}_{#1}\left(#2 \times #3, #4\right)}
\newcommand{\GaussSumCharacter}[3]{\tau\left(#1 \times #2, #3\right)}
\newcommand{\ladicnumbers}{\algebraicClosure{\mathbb{Q}_{\ell}}}
\newcommand{\IsometryGroup}{\mathrm{Isom}}
\newcommand{\lieAlgebra}{\mathfrak{g}}
\newcommand{\DeligneLusztigInduction}[2]{\mathrm{R}_{#1}^{#2}}
\newcommand{\algebraicGroup}[1]{\boldsymbol{\mathrm{#1}}}
\newcommand{\LusztigSeries}[2]{\mathcal{E}\left(#1, (#2)\right)}
\newcommand{\characteristicPolynomial}{\operatorname{CharPoly}}
\newcommand{\DualFrobeniusFixedPoints}[2][\Frobenius^{\ast}]{\algebraicGroup{#2}^{\ast #1}}
\newcommand{\FrobeniusFixedPoints}[2][\Frobenius]{\algebraicGroup{#2}^{#1}}
\newcommand{\CharacterLattice}[1]{X^{\ast}\left(#1\right)}
\newcommand{\CocharacterLattice}[1]{X_{\ast}\left(#1\right)}

@@

We define a normalization factor as follows $$c_{\hermitianSpace}\left(\chi, \fieldCharacter\right) = \begin{dcases}
	\tau\left(\chi^2, \fieldCharacter\right)^{\frac{\dim_{\finiteField} \hermitianSpace}{2}} & \quadraticExtension = \finiteField\\
	\tau\left(\chi\restriction_{\multiplicativegroup{\finiteField}}, \fieldCharacter\right)^{\dim_{\quadraticExtension} \hermitianSpace} & \quadraticExtension \ne \finiteField
\end{dcases}.$$
Notice that if $\quadraticExtension \ne \finiteField$ then by the Hasse--Davenport relation $\tau\left(\chi\restriction_{\multiplicativegroup{\finiteField}}, \fieldCharacter\right)^{2} = \tau\left(\involutionPlusOne{\chi}, \fieldCharacter_{\quadraticExtension}\right)$.

For an irreducible representation  $\tau$ of $\GL_k\left(\quadraticExtension\right)$, define $$\dblGammaFactor{\tau}{\chi}{\fieldCharacter} = \GaussSumScalar{\tau \times \minusInvolution{\chi}}{\fieldCharacter} \GaussSumScalar{\Contragradient{\tau} \times \chi^{-1}}{\fieldCharacter}.$$
For an irreducible representation $\pi$ of $G$, define the \emph{doubling method gamma factor} $$\dblGammaFactorSpace{\hermitianSpace}{\pi}{\chi}{\fieldCharacter} = \centralCharacter{\pi}\left(-1\right) \frac{\dblJacobiSumScalar{\pi}{\chi}}{c_{\hermitianSpace}\left(\chi, \fieldCharacter\right)}.$$
Using this notation and using the multiplicativity property of Kondo's Gauss sum and \Cref{thm:multiplicativity-in-terms-of-gauss-sums}, we have the following multiplicativity property.

@@

\begin{remark}
	The ``correct'' gamma factor (expected by Langlands functoriality) is given by $$\dblLanglandsGammaFactorSpace{\hermitianSpace}{\pi}{\chi}{\fieldCharacter} = \dblGammaFactorSpace{\hermitianSpace}{\pi}{\chi}{\fieldCharacter} \cdot \begin{dcases}
				\left(-1\right)^n & \quadraticExtension = \finiteField \text{ and } \epsilon_{\hermitianSpace} = 1\\
				\left(-1\right)^{n+1} \tau\left(\chi^{-1}, \fieldCharacter\right) & \quadraticExtension = \finiteField \text{ and } \epsilon_{\hermitianSpace} = -1 \\
				1 & \text{otherwise.}
	\end{dcases}.$$
	However, we will not need this version for our purposes.
\end{remark}

@@

It is clear that if $\pi$ is an irreducible representation of $H$ then $$\dblVirtualJacobiSumScalar{\trace \pi}{\chi} = \dblJacobiSumScalar{\pi}{\chi} \cdot \dim \pi.$$
Notice that the assignment $g \mapsto \Phi_{\chi}\left(g\right)$ only depends on the semisimple part of $g \in H$. By \cite[Theorem in Section 1.2]{SaitoShinoda2000}, we have the following result.
\begin{proposition}\label{prop:reduction-of-gauss-sum-to-torus}
	For any $\Frobenius$-stable maximal torus $\algebraicGroup{T}$ of $\algebraicGroup{H}$, and any character $\theta \colon T \to \multiplicativegroup{\cComplex}$, where $T = \algebraicGroup{T}^{\Frobenius}$, we have
	$$ \dblVirtualJacobiSumScalar{R_{T, \theta}}{\chi} = \frac{\grpIndex{H}{T}}{\sqrt{\sizeof{\lieAlgebra}}} \sum_{t \in T} \theta\left(t\right) \genJacobiKernel{\chi}\left(t\right).$$
\end{proposition}
In the special case that $\theta$ is in general position, the virtual character $\varepsilon_{\algebraicGroup{H}} \varepsilon_{\algebraicGroup{T}} R_{T, \theta}$ equals $\trace \pi$ for some irreducible representation $\pi$ of $H$ and we have that for this $\pi$,
$$\dblJacobiSumScalar{\pi}{\chi} = \varepsilon_{\algebraicGroup{H}} \varepsilon_{\algebraicGroup{T}} \frac{\sizeof{H}_p}{\sqrt{\sizeof{\lieAlgebra}}} \sum_{t \in T} \theta\left(t\right) \genJacobiKernel{\chi}\left(t\right).$$
Here $\varepsilon_{\algebraicGroup{H}} = \left(-1\right)^{\mathrm{rel.rank} \algebraicGroup{H}}$ and $\varepsilon_{\algebraicGroup{T}} = \left(-1\right)^{\mathrm{rel.rank} \algebraicGroup{T}}$, and $\sizeof{H}_p$ is the size of the $p$-Sylow group of $H$, where $p$ is the characteristic of $\finiteField$.

@@

\subsubsection{Split torus computation II}
We use the results from the previous section to compute the doubling method Jacobi sum for a torus of the form $\multiplicativegroup{\finiteFieldExtension{k}}$. Let $\alpha \colon \multiplicativegroup{\finiteFieldExtension{k}} \to \multiplicativegroup{\cComplex}$ and $\chi \colon \multiplicativegroup{\finiteField} \to \multiplicativegroup{\cComplex}$ be characters such that $\involutionPlusOne{\chi} \ne 1$. Our goal is to compute $$\sum_{\substack{x \in \multiplicativegroup{\finiteFieldExtension{k}}\\
		x \ne -1}} \alpha \left(x\right) \chi\left(\FieldNorm{k}{1}\left(1 + x\right)\right) \chi\left(\FieldNorm{k}{1}\left(1 + \minusInvolution{x}\right)\right).$$
This can be rewritten as
$$\sum_{\substack{x \in \multiplicativegroup{\finiteFieldExtension{k}}\\
		x \ne -1}} \alpha \left(x\right) \minusInvolution{\chi}\left(\FieldNorm{k}{1}\left(x\right)\right) \involutionPlusOne{\chi}\left(\FieldNorm{k}{1}\left(1 + x\right)\right).$$
By the previous section, this equals
\begin{equation*}
	-q^{\frac{k}{2}} \GaussSumCharacter{\alpha^{-1}}{\chi^{-1}}{\fieldCharacter_{\quadraticFieldExtension{k}}} \GaussSumCharacter{\alpha}{\minusInvolution{\chi}}{\fieldCharacter_{\quadraticFieldExtension{k}}} \tau\left(\involutionPlusOne{\chi}, \fieldCharacter_{\quadraticExtension}\right)^k,
\end{equation*}
which equals
\begin{equation}\label{eq:identity-for-multiplicative-group-of-f-k-gauss-sum}
	-q^{\frac{k}{2}} \GaussSumCharacter{\alpha^{-1}}{\chi^{-1}}{\fieldCharacter_{\quadraticFieldExtension{k}}} \GaussSumCharacter{\involution{\alpha}}{\chi^{-1}}{\fieldCharacter_{\quadraticFieldExtension{k}}} \tau\left(\involutionPlusOne{\chi}, \fieldCharacter_{\quadraticExtension}\right)^k,
\end{equation}

Changing variables $x \mapsto \pm x$, we obtain the following.

\begin{theorem}We have the following identity:
\begin{align*}
	& \sum_{\substack{x \in \multiplicativegroup{\finiteFieldExtension{k}}\\
			x \ne \mp 1}} \alpha \left(x\right) \chi\left(\FieldNorm{k}{1}\left(1 \pm x\right)\right) \chi\left(\FieldNorm{k}{1}\left(1 \pm \minusInvolution{x}\right)\right) \\
	=& -\alpha\left(\pm 1\right) q^{\frac{k}{2}} \GaussSumCharacter{\alpha^{-1}}{\chi^{-1}}{\fieldCharacter_{\quadraticFieldExtension{k}}} \GaussSumCharacter{\involution{\alpha}}{\chi^{-1}}{\fieldCharacter_{\quadraticFieldExtension{k}}} \tau\left(\involutionPlusOne{\chi}, \fieldCharacter_{\quadraticExtension}\right)^k.
\end{align*}
\end{theorem}

@@

To summarize, we have
\begin{align*}
	& \sum_{\substack{x \in \NormOneGroup{2m}\\
			x \ne -1}} \theta \left(x\right) \chi\left(\aFieldNorm_{\finiteFieldExtension{2m} \slash \finiteField}\left(1 + x\right)\right) \\
		=& q^{\frac{m}{2}} \tau\left(\chi^{-1}, \fieldCharacter\right)^{2m} \tau\left(\chi, \fieldCharacter\right)^{2m} \tau\left(\chi^{2}, \fieldCharacter\right)^m \tau\left(\theta' \times \chi^{-1}, \fieldCharacter_{2m}\right),
\end{align*}
where $\theta'\left(t\right) = \theta\left(t^{1-q^m}\right)$.
Since $\tau\left(\chi^{-1}, \fieldCharacter\right) = \chi\left(-1\right) \conjugate{\tau\left(\chi, \fieldCharacter\right)}$ and since $\chi \ne 1$, this implies the identity
$$\sum_{\substack{x \in \NormOneGroup{2m}\\
		x \ne -1}} \theta \left(x\right) \chi\left(\aFieldNorm_{\finiteFieldExtension{2m} \slash \finiteField}\left(1 + x\right)\right) = q^{\frac{m}{2}} \tau\left(\chi^{2}, \fieldCharacter\right)^m \tau\left(\theta' \times \chi^{-1}, \fieldCharacter_{2m}\right).$$
Changing variables $x \mapsto \pm x$, we obtain the following. \begin{theorem}
	We have the identity $$\sum_{\substack{x \in \NormOneGroup{2m}\\
			x \ne \mp 1}} \theta \left(x\right) \chi\left(\aFieldNorm_{\finiteFieldExtension{2m} \slash \finiteField}\left(1 \pm x\right)\right) = \theta\left(\pm 1\right) q^{\frac{m}{2}} \tau\left(\chi^{2}, \fieldCharacter\right)^m \tau\left(\theta' \times \chi^{-1}, \fieldCharacter_{2m}\right).$$
\end{theorem}

@@

Changing variables $x \mapsto \pm x$, we obtain the following.
\begin{theorem}
	We have the following identity $$\sum_{\substack{x \in \NormOneGroup{2m}\\
			x \ne -1}} \theta \left(x\right) \chi\left(\aFieldNorm_{\finiteFieldExtension{2m} \slash \finiteFieldExtension{2}}\left(1 + x\right)\right) = \theta\left(\pm 1\right) q^{\frac{m}{2}} \tau\left(\chi \restriction_{\multiplicativegroup{\finiteField}}, \fieldCharacter\right)^m \GaussSumCharacter{\left(\theta'\right)^{-1}}{\chi^{-1}}{\fieldCharacter_{2m}}$$
\end{theorem}

@@

\begin{theorem}\label{thm:computation-of-doubling-gauss-sum-scalar-for-deligne-lusztig-characters}
	Suppose that $T \cong \prod_{j=1}^r \multiplicativegroup{\quadraticFieldExtension{k_j}} \times \prod_{i=1}^s \finiteFieldExtension{2m_i}^1$ is a maximal torus of $H$. Suppose that $\theta \colon T \to \multiplicativegroup{\cComplex}$ is a character and that under the above isomorphism $\theta = \alpha_1 \times \dots \times \alpha_r \times \theta_1 \times \dots \times \theta_s$, where $\alpha_j \colon \multiplicativegroup{\quadraticFieldExtension{k_j}} \to \multiplicativegroup{\cComplex}$ and $\theta_i \colon \finiteFieldExtension{2m_i}^1 \to \multiplicativegroup{\cComplex}$ are characters for every $i$ and $j$. Denote \begin{align*}
		g_T\left(\chi, \theta, \fieldCharacter\right)
		= & \prod_{i=1}^s \GaussSumCharacter{\left(\theta'_i\right)^{-1}}{\chi^{-1}}{\fieldCharacter_{2m_i}} \cdot \prod_{j=1}^r \GaussSumCharacter{\involution{\alpha_j}}{\chi^{-1}}{\fieldCharacter_{\quadraticFieldExtension{k_j}}} \GaussSumCharacter{\alpha_j^{-1}}{\chi^{-1}}{\fieldCharacter_{\quadraticFieldExtension{k_j}}}.
	\end{align*} Then
	\begin{align*}
		 \dblVirtualJacobiSumScalar{R_{T, \theta}}{\chi} = \varepsilon_{\algebraicGroup{H}} R_{T,\theta}\left(1\right) c_{\hermitianSpace}\left(\chi, \fieldCharacter\right) g_T\left(\chi, \theta, \fieldCharacter\right) \prod_{i=1}^s \theta_i\left(\pm 1\right) \cdot \prod_{j=1}^r \alpha_j\left(\pm 1\right).
	\end{align*}
	\elad{Modify this so it is also true in the odd special orthogonal group case: it is zero for $-$ and there are factors $\theta\left(2\right)$ for $+$}
\end{theorem}
\begin{proof}
	This follows from the results mentioned above and from the formula $$R_T\left(1\right) = \varepsilon_{\algebraicGroup{H}} \varepsilon_{\algebraicGroup{T}} \frac{\grpIndex{H}{T}}{\sizeof{H}_p},$$ the fact that $$ \sizeof{\lieAlgebra} = \sizeof{H}_p^2 \cdot q^{\frac{\dim_{\finiteField} \hermitianSpace}{2}},$$
	and the fact \begin{equation*}
		\varepsilon_{\algebraicGroup{T}} = \left(-1\right)^r.
	\end{equation*}
\end{proof}
\begin{remark}
	We have $$\varepsilon_{\algebraicGroup{G}} = \left(-1\right)^{\frac{\dim_{\finiteField} \hermitianSpace}{2}}.$$ If $\finiteField = \quadraticExtension$ then we also have $$\varepsilon_{\algebraicGroup{\GroupExtension{G}}} = \left(-1\right)^{\frac{\dim_{\finiteField} \hermitianSpace}{2} + 1}.$$
\end{remark}
\begin{remark}
	Suppose that $\pi$ is any irreducible representation of $H$ such that $\trace \pi$ appears with non-zero coefficient in $R_{T,\theta}$. Then the central character of $\pi$ satisfies
	\begin{align*}
		\centralCharacter{\pi}\left(-1\right) = \theta\left(-1\right) =& \prod_{j=1}^r \alpha_j\left(-1\right) \cdot \prod_{i=1}^s \theta_i\left(-1\right).
	\end{align*}
	Thus \begin{align*}
		\dblVirtualJacobiSumScalar{R_{T, \theta}}{\chi} =& R_{T,\theta}\left(1\right) c_{\hermitianSpace}\left(\chi, \fieldCharacter\right) \centralCharacter{\pi}\left(-1\right) g_T\left(\chi, \theta, \fieldCharacter\right).
	\end{align*}
\end{remark}

>>>>

\newcommand{\GaussSumScalar}[2]{\mathrm{G}\left(#1, #2\right)}
\newcommand{\dblJacobiSumScalar}[2]{\mathrm{J}_{\pm}^{\mathrm{dbl}}\left(#1, #2\right)}
\newcommand{\posDblJacobiSumScalar}[2]{\mathrm{J}_{+}^{\mathrm{dbl}}\left(#1, #2\right)}
\newcommand{\negDblJacobiSumScalar}[2]{\mathrm{J}_{-}^{\mathrm{dbl}}\left(#1, #2\right)}
\newcommand{\dblVirtualJacobiSumScalar}[2]{\mathrm{j}_{\pm}^{\mathrm{dbl}}\left(#1, #2\right)}
\newcommand{\dblPosVirtualJacobiSumScalar}[2]{\mathrm{j}_{+}^{\mathrm{dbl}}\left(#1, #2\right)}
\newcommand{\dblNegVirtualJacobiSumScalar}[2]{\mathrm{j}_{-}^{\mathrm{dbl}}\left(#1, #2\right)}
\newcommand{\posDblVirtualJacobiSumScalar}[2]{\mathrm{j}_{+}^{\mathrm{dbl}}\left(#1, #2\right)}
\newcommand{\dblGammaFactor}[3]{\Gamma^{\mathrm{dbl}}\left(#1 \times #2, #3\right)}
\newcommand{\dblGammaFactorSpace}[4]{\Gamma^{\mathrm{dbl}}_{#1}\left(#2 \times #3, #4\right)}
\newcommand{\dblLanglandsGammaFactorSpace}[4]{\gamma^{\mathrm{dbl}}_{#1}\left(#2 \times #3, #4\right)}
\newcommand{\GaussSumCharacter}[3]{\tau\left(#1 \times #2, #3\right)}
\newcommand{\ladicnumbers}{\algebraicClosure{\mathbb{Q}_{\ell}}}
\newcommand{\IsometryGroup}{\mathrm{Isom}}
\newcommand{\lieAlgebra}{\mathfrak{g}}
\newcommand{\DeligneLusztigInduction}[2]{\mathrm{R}_{#1}^{#2}}
\newcommand{\algebraicGroup}[1]{\boldsymbol{\mathrm{#1}}}
\newcommand{\LusztigSeries}[2]{\mathcal{E}\left(#1, (#2)\right)}
\newcommand{\characteristicPolynomial}{\operatorname{CharPoly}}
\newcommand{\DualFrobeniusFixedPoints}[2][\Frobenius^{\ast}]{\algebraicGroup{#2}^{\ast #1}}
\newcommand{\FrobeniusFixedPoints}[2][\Frobenius]{\algebraicGroup{#2}^{#1}}
\newcommand{\CharacterLattice}[1]{X^{\ast}\left(#1\right)}
\newcommand{\CocharacterLattice}[1]{X_{\ast}\left(#1\right)}

@@

We define a normalization factor as follows $$c_{\hermitianSpace}\left(\chi, \fieldCharacter\right) = \begin{dcases}
	\tau\left(\chi^2, \fieldCharacter\right)^{\frac{\dim_{\finiteField} \hermitianSpace}{2}} & \quadraticExtension = \finiteField,\,\, \dim_{\finiteField} \hermitianSpace \text{ is even,}\\
	\chi\left(2\right) \tau\left(\chi^2, \fieldCharacter\right)^{\frac{\dim_{\finiteField} \hermitianSpace - 1}{2}} & \quadraticExtension = \finiteField,\,\, \dim_{\finiteField} \hermitianSpace \text{ is odd,}	\\
	\tau\left(\chi\restriction_{\multiplicativegroup{\finiteField}}, \fieldCharacter\right)^{\dim_{\quadraticExtension} \hermitianSpace} & \quadraticExtension \ne \finiteField.
\end{dcases}$$
Notice that if $\quadraticExtension \ne \finiteField$ then by the Hasse--Davenport relation $\tau\left(\chi\restriction_{\multiplicativegroup{\finiteField}}, \fieldCharacter\right)^{2} = \tau\left(\involutionPlusOne{\chi}, \fieldCharacter_{\quadraticExtension}\right)$.

For an irreducible representation  $\tau$ of $\GL_k\left(\quadraticExtension\right)$, define $$\dblGammaFactor{\tau}{\chi}{\fieldCharacter} = \GaussSumScalar{\tau \times \minusInvolution{\chi}}{\fieldCharacter} \GaussSumScalar{\Contragradient{\tau} \times \chi^{-1}}{\fieldCharacter}.$$
For an irreducible representation $\pi$ of $G$, define the \emph{doubling method gamma factor} $$\dblGammaFactorSpace{\hermitianSpace}{\pi}{\chi}{\fieldCharacter} = \frac{\posDblJacobiSumScalar{\pi}{\chi}}{c_{\hermitianSpace}\left(\chi, \fieldCharacter\right)}.$$
Using this notation and using the multiplicativity property of Kondo's Gauss sum and \Cref{thm:multiplicativity-in-terms-of-gauss-sums}, we have the following multiplicativity property.

@@

\begin{remark}
	The ``correct'' gamma factor (expected by Langlands functoriality) is given by $$\dblLanglandsGammaFactorSpace{\hermitianSpace}{\pi}{\chi}{\fieldCharacter} = \dblGammaFactorSpace{\hermitianSpace}{\pi}{\chi}{\fieldCharacter} \cdot \begin{dcases}
				\left(-1\right)^n & \quadraticExtension = \finiteField \text{ and } \epsilon_{\hermitianSpace} = 1\\
				\left(-1\right)^{n+1} \tau\left(\chi^{-1}, \fieldCharacter\right) & \quadraticExtension = \finiteField \text{ and } \epsilon_{\hermitianSpace} = -1 \\
				1 & \text{otherwise.}
	\end{dcases}.$$
	However, we will not need this version for our purposes. \elad{Check odd orthogonal case. Probably needs a sign modification (the formula needs to be $\left(-1\right)^{\text{dual group dimension}} \times \text{product of Gauss sums}$). $\varepsilon_{\algebraicGroup{G}}$ contributes half of the dimension so that's why we add another half here.}
\end{remark}

@@

It is clear that if $\pi$ is an irreducible representation of $H$ then $$\dblVirtualJacobiSumScalar{\trace \pi}{\chi} = \dblJacobiSumScalar{\pi}{\chi} \cdot \dim \pi.$$
Notice that the assignments $g \mapsto \posJacobiKernel{\chi}\left(g\right)$ and $g \mapsto \negJacobiKernel{\chi}\left(g\right)$ only depends on the semisimple part of $g \in H$. By \cite[Theorem in Section 1.2]{SaitoShinoda2000}, we have the following result.
\begin{proposition}\label{prop:reduction-of-gauss-sum-to-torus}
	For any $\Frobenius$-stable maximal torus $\algebraicGroup{T}$ of $\algebraicGroup{H}$, and any character $\theta \colon T \to \multiplicativegroup{\cComplex}$, where $T = \algebraicGroup{T}^{\Frobenius}$, we have
	$$ \dblVirtualJacobiSumScalar{R_{T, \theta}}{\chi} = \frac{\grpIndex{H}{T}}{\sqrt{\sizeof{\lieAlgebra}}} \sum_{t \in T} \theta\left(t\right) \genJacobiKernel{\chi}\left(t\right).$$
\end{proposition}
In the special case that $\theta$ is in general position, the virtual character $\varepsilon_{\algebraicGroup{H}} \varepsilon_{\algebraicGroup{T}} R_{T, \theta}$ equals $\trace \pi$ for some irreducible representation $\pi$ of $H$ and we have that for this $\pi$,
$$\dblJacobiSumScalar{\pi}{\chi} = \varepsilon_{\algebraicGroup{H}} \varepsilon_{\algebraicGroup{T}} \frac{\sizeof{H}_p}{\sqrt{\sizeof{\lieAlgebra}}} \sum_{t \in T} \theta\left(t\right) \genJacobiKernel{\chi}\left(t\right).$$
Here $\varepsilon_{\algebraicGroup{H}} = \left(-1\right)^{\mathrm{rel.rank} \algebraicGroup{H}}$ and $\varepsilon_{\algebraicGroup{T}} = \left(-1\right)^{\mathrm{rel.rank} \algebraicGroup{T}}$, and $\sizeof{H}_p$ is the size of the $p$-Sylow group of $H$, where $p$ is the characteristic of $\finiteField$.

@@

\subsubsection{Split torus computation II}
We use the results from the previous section to compute the doubling method Jacobi sum for a torus of the form $\multiplicativegroup{\finiteFieldExtension{k}}$. Let $\alpha \colon \multiplicativegroup{\finiteFieldExtension{k}} \to \multiplicativegroup{\cComplex}$ and $\chi \colon \multiplicativegroup{\finiteField} \to \multiplicativegroup{\cComplex}$ be characters such that $\involutionPlusOne{\chi} \ne 1$. Our goal is to compute $$\sum_{\substack{x \in \multiplicativegroup{\finiteFieldExtension{k}}\\
		x \ne -1}} \alpha \left(x\right) \chi\left(\FieldNorm{k}{1}\left(1 + x\right)\right) \chi\left(\FieldNorm{k}{1}\left(1 + \minusInvolution{x}\right)\right).$$
This can be rewritten as
$$\sum_{\substack{x \in \multiplicativegroup{\finiteFieldExtension{k}}\\
		x \ne -1}} \alpha \left(x\right) \minusInvolution{\chi}\left(\FieldNorm{k}{1}\left(x\right)\right) \involutionPlusOne{\chi}\left(\FieldNorm{k}{1}\left(1 + x\right)\right).$$
By the previous section, this equals
\begin{equation*}
	-q^{\frac{k}{2}} \GaussSumCharacter{\alpha^{-1}}{\chi^{-1}}{\fieldCharacter_{\quadraticFieldExtension{k}}} \GaussSumCharacter{\alpha}{\minusInvolution{\chi}}{\fieldCharacter_{\quadraticFieldExtension{k}}} \tau\left(\involutionPlusOne{\chi}, \fieldCharacter_{\quadraticExtension}\right)^k,
\end{equation*}
which equals
\begin{equation*}
	-q^{\frac{k}{2}} \GaussSumCharacter{\alpha^{-1}}{\chi^{-1}}{\fieldCharacter_{\quadraticFieldExtension{k}}} \GaussSumCharacter{\involution{\alpha}}{\chi^{-1}}{\fieldCharacter_{\quadraticFieldExtension{k}}} \tau\left(\involutionPlusOne{\chi}, \fieldCharacter_{\quadraticExtension}\right)^k,
\end{equation*}
Thus we obtained the following.
\begin{theorem}We have the identity
\begin{align*}
	& \sum_{\substack{x \in \multiplicativegroup{\finiteFieldExtension{k}}\\
			x \ne -1}} \alpha \left(x\right) \chi\left(\FieldNorm{k}{1}\left(1 + x\right)\right) \chi\left(\FieldNorm{k}{1}\left(1 + \minusInvolution{x}\right)\right) \\
	=& -q^{\frac{k}{2}} \GaussSumCharacter{\alpha^{-1}}{\chi^{-1}}{\fieldCharacter_{\quadraticFieldExtension{k}}} \GaussSumCharacter{\involution{\alpha}}{\chi^{-1}}{\fieldCharacter_{\quadraticFieldExtension{k}}} \tau\left(\involutionPlusOne{\chi}, \fieldCharacter_{\quadraticExtension}\right)^k.
\end{align*}
\end{theorem}

@@

To summarize, we have
\begin{align*}
	& \sum_{\substack{x \in \NormOneGroup{2m}\\
			x \ne -1}} \theta \left(x\right) \chi\left(\aFieldNorm_{\finiteFieldExtension{2m} \slash \finiteField}\left(1 + x\right)\right) \\
		=& q^{\frac{m}{2}} \tau\left(\chi^{-1}, \fieldCharacter\right)^{2m} \tau\left(\chi, \fieldCharacter\right)^{2m} \tau\left(\chi^{2}, \fieldCharacter\right)^m \tau\left(\theta' \times \chi^{-1}, \fieldCharacter_{2m}\right),
\end{align*}
where $\theta'\left(t\right) = \theta\left(t^{1-q^m}\right)$.
Since $\tau\left(\chi^{-1}, \fieldCharacter\right) = \chi\left(-1\right) \conjugate{\tau\left(\chi, \fieldCharacter\right)}$ and since $\chi \ne 1$, this implies the following. \begin{theorem}
	We have the identity $$\sum_{\substack{x \in \NormOneGroup{2m}\\
			x \ne -1}} \theta \left(x\right) \chi\left(\aFieldNorm_{\finiteFieldExtension{2m} \slash \finiteField}\left(1 + x\right)\right) = q^{\frac{m}{2}} \tau\left(\chi^{2}, \fieldCharacter\right)^m \tau\left(\theta' \times \chi^{-1}, \fieldCharacter_{2m}\right).$$
\end{theorem}

@@

Thus, we obtained the following.
\begin{theorem}
	We have the following identity $$\sum_{\substack{x \in \NormOneGroup{2m}\\
			x \ne -1}} \theta \left(x\right) \chi\left(\aFieldNorm_{\finiteFieldExtension{2m} \slash \finiteFieldExtension{2}}\left(1 + x\right)\right) = q^{\frac{m}{2}} \tau\left(\chi \restriction_{\multiplicativegroup{\finiteField}}, \fieldCharacter\right)^m \GaussSumCharacter{\left(\theta'\right)^{-1}}{\chi^{-1}}{\fieldCharacter_{2m}}$$
\end{theorem}

@@

\begin{theorem}\label{thm:computation-of-doubling-gauss-sum-scalar-for-deligne-lusztig-characters}
	Suppose that $T \cong \prod_{j=1}^r \multiplicativegroup{\quadraticFieldExtension{k_j}} \times \prod_{i=1}^s \finiteFieldExtension{2m_i}^1$ is a maximal torus of $H$. Suppose that $\theta \colon T \to \multiplicativegroup{\cComplex}$ is a character and that under the above isomorphism $\theta = \alpha_1 \times \dots \times \alpha_r \times \theta_1 \times \dots \times \theta_s$, where $\alpha_j \colon \multiplicativegroup{\quadraticFieldExtension{k_j}} \to \multiplicativegroup{\cComplex}$ and $\theta_i \colon \finiteFieldExtension{2m_i}^1 \to \multiplicativegroup{\cComplex}$ are characters for every $i$ and $j$. Denote \begin{align*}
		g_T\left(\chi, \theta, \fieldCharacter\right)
		= & \prod_{i=1}^s \GaussSumCharacter{\left(\theta'_i\right)^{-1}}{\chi^{-1}}{\fieldCharacter_{2m_i}} \cdot \prod_{j=1}^r \GaussSumCharacter{\involution{\alpha_j}}{\chi^{-1}}{\fieldCharacter_{\quadraticFieldExtension{k_j}}} \GaussSumCharacter{\alpha_j^{-1}}{\chi^{-1}}{\fieldCharacter_{\quadraticFieldExtension{k_j}}}.
	\end{align*} Then
	\begin{align*}
		 \dblPosVirtualJacobiSumScalar{R_{T, \theta}}{\chi} = \varepsilon_{\algebraicGroup{H}} R_{T,\theta}\left(1\right) c_{\hermitianSpace}\left(\chi, \fieldCharacter\right) g_T\left(\chi, \theta, \fieldCharacter\right),
	\end{align*}
	and \begin{align*}
		\dblNegVirtualJacobiSumScalar{R_{T, \theta}}{\chi} = \begin{dcases}
			0 & \dim_{\finiteField} \hermitianSpace \text{ is odd.}\\
			\dblPosVirtualJacobiSumScalar{R_{T, \theta}}{\chi} \cdot \prod_{i=1}^s \theta_i\left(-1\right) \cdot \prod_{j=1}^r \alpha_j\left(-1\right) & \text{ otherwise.}
		\end{dcases}
	\end{align*}
\end{theorem}
\begin{proof}
	Notice that if $\dim_{\finiteField} \hermitianSpace$ is odd then $\negJacobiKernel{\chi}\left(t\right) = 0$ for any $t \in T$, as $\idmap_{\hermitianSpace} - t$ always has determinant zero.

	We have that if $t$ corresponds to $\left(a_1,\dots,a_r,t_1,\dots,t_s\right) \in \prod_{j=1}^r \multiplicativegroup{\quadraticFieldExtension{k_j}} \times \prod_{i=1}^s \NormOneGroup{2m_i}$ such that $\idmap_{\hermitianSpace} + t$ is invertible then \begin{align*}
		\posJacobiKernel{\chi}\left(t\right) =& \prod_{i=1}^s \chi\left(\FieldNorm{2m_i}{1}\left(1 + t_i\right)\right) \cdot \prod_{j=1}^r \chi\left(\aFieldNorm_{\quadraticFieldExtension{k_j} / \quadraticExtension}\left(1 + a_j\right)\right) \chi\left(\aFieldNorm_{\quadraticFieldExtension{k_j} / \quadraticExtension}\left(1 + \minusInvolution{a_j}\right)\right) \\
		& \times  \begin{dcases}
			\chi\left(2\right) & \dim_{\finiteField} \hermitianSpace \text{ is odd},\\
			1 & \text{otherwise}
		\end{dcases}.
	\end{align*}
	If $t$ corresponds to $\left(a_1,\dots,a_r,t_1,\dots,t_s\right) \in \prod_{j=1}^r \multiplicativegroup{\quadraticFieldExtension{k_j}} \times \prod_{i=1}^s \NormOneGroup{2m_i}$ such that $\idmap_{\hermitianSpace} - t$ is invertible then \begin{align*}
		\negJacobiKernel{\chi}\left(t\right) =& \begin{dcases}
			0 & \dim_{\finiteField} \hermitianSpace \text{ is odd},\\
			\posJacobiKernel{\chi}\left(-t\right) & \text{otherwise.}
		\end{dcases}
	\end{align*}
